---
title: 'Lab 5: Exploratory Analytics'
author: "Loura Shiny M"
date: "10/09/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Exploratory Analytics in R

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to learn and execute the different concepts 
under Exploratory Analytics in R by practicing different questions.

## To study / implement:

From this Lab exercise we will be learning about Exploratory Analytics.


## Dataset Description:

The dataset used is from package MASS. The Student Survey Dataset is pool of 
information on various deatails collected from students at the University of 
Adelaide. It contains information such as the sex of the student, their height, 
age, pulse, which hand they use to write, fold, clap,etc. 

### Name of the Dataset: 

Student Survey Dataset

------------------------------------------------------------------------

## Questions


Loading in the dataset:

```{r}
library(MASS)
```

```{r}
library(tidyverse)
```

```{r}
str(survey)
```

```{r}
survey=na.omit(survey)
```


1. Filter out the females students who clap with their left hand on top. 

```{r}
x=survey%>% filter(Sex=="Female" & Clap=="Left")
x
```

2. Get the details of students who don't exercise at all.

```{r}
filter(survey,Exer=="None")
```

3. Plot the height of the students.

```{r}
plot(survey$Height,type="b",ylab="Height")
```

4. Plot the age against the height of the students.

```{r}
plot(survey$Height,survey$Age, xlab="Height",ylab="Age")
```

5. Plot a histogram to view the pulse frequency of the different students.

```{r}
hist(survey$Pulse,main="Pulse",xlab="Pulse")
```

6.Show the differnce in the number of left handede and right handed students 
using a bar graph.

```{r}
x=table(survey$W.Hnd)
barplot(x,main="Left and Right Handed  Students", xlab="WritingHand", ylab="Value",col=c("Blue","Pink"),legend = rownames(x))
```


7. Box plot the age of the students.

```{r}
boxplot(survey$Age,main="Age")
```

8. Show the different type of somkers using bar graph.

```{r}
x=table(survey$Smoke)
barplot(x,main="Type of Smokers",xlab="Smoking Frequency",ylab="Number",col=c("Blue","Pink","Yellow","Green"),legend = rownames(x))
```

9.Plot a histogram of non-writing hand of the students.

```{r}
hist(survey$NW.Hnd,mai="Non-Writing Hand",xlab="Values")
```


10.Show in a pie chart on how many students gave in their heights in metrics and 
how many in imperial.

```{r}
x=survey %>% filter(Sex=="Male" & W.Hnd=="Left")
y=table(x$M.I)
pie(y,col=c("red","green"),main="M.I")
```


* * *
## Conclusion:
Using the dataset of The Student Survey, we learned to do Exploratory Analytics
by using various concepts of data visualization.

