---
title: 'Lab 3: Data Manipulation in R'
author: "Loura Shiny M"
date: "19/08/2021"
output:
  html_document: default
  pdf_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Data Manipulation in R

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to learn and execute the different concepts 
under data manipulation in R by doing the instructed questions.

## To study / implement:

From this Lab exercise we will be learning about data manipulation in R.


## Dataset Description

For this experiment we will be using the flight dataset from the package 
nycflights13.

### Name of the Dataset: 

Flight

### Description of the fields : 

The data set contains the details of different flights such as day on which it 
was boarded, it's departure time, scheduled departure time, origin, destination,
air time, carrier, flight number, timings, etc.


```{r}
library("nycflights13")
```


```{r}
View(flights)
str(flights)
```
```{r}
library(dplyr)
```

------------------------------------------------------------------------

## Questions


1. Find all flights that had an arrival delay of two or more hours.

```{r}
filter(flights, (arr_delay/60)>=2)
```

2. Find all flights that had arrived more than two hours late, but didn’t leave 
late. 

```{r}
filter(flights, (arr_delay/60)>=2, dep_delay<=0)
```

3. How many flights have a missing dep_time? 

```{r}
sum(is.na(flights$dep_time))
```

4. List the flights that are operated by American Airlines(Carrier is ‘AA’).

```{r}
filter(flights,carrier=="AA")
```

5. List the flights that departed during August 2013.

```{r}
filter(flights,year==2013,month==8)
```

6. Find the most delayed flights considering their arrival time alone.

```{r}
x<-arrange(flights,desc(arr_delay))
head(x)
```


7. List the flights that operate between LGA and IAH airports.

```{r}
filter(flights,origin=="LGA"|origin=="IAH",dest=="IAH"|dest=="LGA")
```

8. List the flights that has the destination of IAH.

```{r}
filter(flights,dest=="IAH")
```

9. Display the details of the flight that covers the longest distance / shortest 
distance. (flight number, carrier, origin, destination and distance)

```{r}
flights%>%filter(distance==max(distance)|distance==min(distance))%>%select(flight,carrier,origin,dest,distance)%>%arrange(distance)
```

10.Which carrier has the worst delays

```{r}
flights %>%group_by(carrier) %>%
summarise(arr_delay = mean(arr_delay, na.rm = TRUE))%>%arrange(desc(arr_delay))
```

* * *
## Conclusion:
Using the dataset flights, we learned about how to manipulate a dataset according to our needs.



