---
title: 'Lab 9: Sports Analysis in R'
author: "Loura Shiny M"
date: "21/10/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Sports Analysis in R

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to perform descriptive and predictive 
analytics from a suitable dataset chosen in sports domain. The dataset can be 
either for regression or classification. 

## To study / implement:

From this Lab exercise we will be learning about Sport Ananlysis in R. 


## Dataset Description:

This dataset Toughesht Sports by Skill ranks the toughest sports on their skill
level by ESPN. They include factors such as endurance, strength, power, speed, 
agiligty, flexibility, etc. 

https://www.kaggle.com/rishidamarla/toughest-sports-by-skill


## Objectives:

The objective her is to perform descriptive and predictive analytics for the 
chosen dataset - Toughest Sports by Skill. The descriptive analysis will give a 
summary of the dataset and answer the questions asked in that section
and using linear regression we will do predictive analysis. 

------------------------------------------------------------------------


## Pre-Processing:

## Sequence of Steps:


### Importing required libraries 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(caret)
library(datarium)
```


### Importing Dataset

```{r}
sports = read.csv("toughestsport.csv")
```


### View structure of dataset 

```{r}
str(sports)
```


### Perform Data Clenaing

```{r}
names(sports)[names(sports) == 'ï..SPORT'] <- "SPORT"
```

Here we'll change the name of the column from "ï..SPORT" to "SPORT" for 
readability purpose.


Structure of dataset

```{r}
str(sports)
```


### Removing missing values 

```{r}
sports = na.omit(sports)
```


### Summary of the dataset 

```{r}
summary(sports)
```


## Descriptive Analysis:

##### Range of Durability

```{r}
range(sports$DUR)
```

The range od the Durability ranges from 0.75 to 8.50


##### Histogram of Endurance

```{r}
hist(sports$END,main="Endurance in Sports",xlab="Endurance")
```


##### Division of how much power is required to play the sports in the first 5 ranks 

```{r}
x=sports %>% filter(RANK %in% c(1,2,3,4,5))
y=table(x$PWR)
pie(y,col=c("red","green","blue","yellow","orange"),main="Power")
```


The pie chart (pictorial representation) of the power among the first 5 sports.

##### Boxplot of speed

```{r}
boxplot(sports$SPD,mai="Speed")
```

Box plots shows that there is no outlier detected.


##### Plotting the agility of the sports

```{r}
plot(sports$AGI,type="b",ylab="Agility")
```

As the rank value of agility increases the rank of the sport goes up.


## Predictive Analysis:

### Model Building

Now to perform linear regression we need to remove the SPORT column. Thus, 

```{r}
sports_data=sports[,-c(1)]
str(sports_data)
```


##### Train the data considering TOTAL from the dataset

```{r}
set.seed(123)
train_samples <- sports_data$TOTAL %>%
createDataPartition(p=0.8,list=FALSE)
```


##### Split the data into training and testing data

```{r}
train <- sports_data[train_samples,]
test <- sports_data[-train_samples,]
```


##### Build Linear model for the TOTAL and other columns in the datset. 
We need to see which attribute has best p-values


```{r}
sports_model<- lm(TOTAL~.,data=sports_data)
summary(sports_model)
```


From summary we can see that Adjusted R-squared values is 1. This is good as it 
indicates that a large propotion of the variability in the outcome has been 
explained by the regression model. Plus, 3 *'s (pr lesser than 0.05 ) for the 
columns END, STR, PWR, SPD, AGI, FLX, NER, DUR, HAN, ANA show that we can use 
them to build linear model. 

Now, considering STR (strength) and total fo our model.


##### Outlier Detection

```{r}
boxplot(sports_data$STR,ylab="STR")
```

No, outlier is detected from boxplot.


### Performance analysis:

##### Predicting the test

```{r}
sp<- sports_model%>%predict(test)
```


##### Finding RMSE

```{r}
RMSE <- RMSE(sp,test$TOTAL)
RMSE
```

The RMSE value is near to 0 which implese that a good outcome for the model.


##### Finding R2

```{r}
R2 <- R2(sp,test$TOTAL)
R2
```

The R2 value is great. This means the model output will be better.


##### Correlation Analysis

```{r}
cor(sports_data$STR,sports_data$TOTAL)
```

Positive value indicates that the that there is positive correlation between 
STR and TOTAL, i.e., they are directly  propositional to each other. 


##### Linear Model between TOTAL and STR

```{r}
model <- lm(TOTAL~STR,data=sports_data)
model
```


```{r}
summary(model)
```


```{r}
sigma(model) 
```


```{r}
sigma(model)*100/mean(sports_data$TOTAL)  #percentage error
```

The percentage of error is 18%.


```{r}
confint(model)
```


##### Finding the confidence of STR

```{r}
STR=6.05
sdt <- data.frame(STR)
conf<- predict(model,sdt,level = .95,interval = "confidence")
conf
```


##### Finding the prediction of STR

```{r}
sp <- predict(model,sdt,level = .95,interval = "prediction")
sp
```


##### Plotting model

```{r}
plot(model,1)
```




------------------------------------------------------------------------


## Results and Discussion: 

From descriptive and predictive analysis we got the desired results. 
In descriptive analysis, we can visualize the data from according to one's wish 
using different graphs. Using linear regression, we can predict the values of 
TOTAL from all the attributes as all of them had 3 *'s. And the example with 
strength, STR was shown.


* * *
## Conclusion:
Using the dataset Toughest Sports by Skill we learnt about Descriptive and 
Predictive Analysis in R by performing them on the dataset.


## References:

https://pestleanalysis.com/descriptive-analysis/
https://machinelearningmastery.com/classification-versus-regression-in-machine-learning/
https://machinelearningmastery.com/machine-learning-in-r-step-by-step/
s
