---
title: 'Lab 4: Data Cleaning in R'
author: "Loura Shiny M"
date: "26/08/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Data Cleaning in R

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to learn and execute the different concepts 
under data cleaning in R by doing the instructed questions for any dataset of 
choice.

## To study / implement:

From this Lab exercise we will be learning about data cleaning in R.


## Dataset Description:

The dataset Throw Competitions Top 100 Marks, contains the information about the 
top 100 sport people from each discipline. I have choosen the disciple Discus 
throw under Men. It conatins their rank, mark, competitor's name, day of birth,
nationality, position, venue, date, and result score.

https://www.kaggle.com/ricfraga/throw-competition-data/version/1

### Name of the Dataset: 

Discus Throw Men Throw Competitions Top 100 Marks

### Description of the fields : 

The Discus Throw Men Competitions Top 100 Marks dataset contains the detaisl of 
their rank, mark, competitor's name, day of birth, nationality, position, venue,
date, and result score.

------------------------------------------------------------------------

## Questions

1. Identify a suitable dataset and describe the dataset.

```{r}
throw<-read.csv(file='Discus Throw Men.csv')
```

```{r}
str(throw)
```

The dataset I have choosen is the Throw Competitions Top 100 Marks. Under this 
I have choosen the throw sport Discus for Men. The dataset contains Discus Throw 
Men Competitions Top 100 Marks. it contains details such as their rank, mark, 
competitor's name, day of birth, nationality, position, venue, date of 
competition,and result score.


2. List all the variable/attribute names.

```{r}
names(throw)
```


3. Identify the mislabeled / long variable names and change them into shorter 
names.

```{r}
names(throw)[names(throw) == 'X'] <- "S.No"
```

```{r}
names(throw)
```


4. Check whether your data contains any unique ID as one of the variables.
Remove the duplicates in uniqueID, if present.

```{r}
any(duplicated(throw$S.No))
```

5. Check for any inconsistencies in the string data. Eg. Male/male/m/M. 
Change them to common format using appropriate method.

```{r}
throw[throw$Rank=="1f1"]<-1
```

```{r}
library(dplyr)
filter(throw,Rank=="1f1")
```

As you can see the rows with isconsistent Rank 1f1 is changed to 1.


6. Check for any invalid values in your dataset. Eg. Age of a person 600..

```{r}
summary(throw$Position)
```

Positions cannot be a negative value but here the minimum value for Position is 
mentioned as -1. Now, we'll change these values to 1.

```{r}
throw$Position[which(throw$Position==-1)]<-1
```

```{r}
summary(throw$Position)
```


7. Check for the missing values. Apply appropriate method to fix it and justify 
it. 

```{r}
sapply(throw,function(x)sum(is.na(x)))
```

From the above result we can see that the columns Marks and Result.Score have 
missing values.


a. Removing that variable - Marks and Result.Score are essential variable. 
So, no we cannot remove them.

```{r}
#throw=subset(throw,select=-c(Mark,ResultScore))
```

b. Removing the samples - For the given situation this might be the best 
solution. We can just remove the rows which contain missing data.

```{r}
throw<-na.omit(throw)
```

```{r}
sapply(throw,function(x)sum(is.na(x)))
```


c. Imputing the missing values - There a few ways through which we can input 
missing values. One is by taking the mean or the median and put that way in the
place of the missing value. The other way is to enter the value which has the 
maximum count in categorical data. But this method won't be right for this data 
as here each player will have a unique mark and result score and cannot be 
computed from the mean. If it was done, it would have been done like this:

throw$Mark[is.na(throw$Mark)]<-median(throw$Mark,na.rm = TRUE)
throw(is.na(throw$Mark))

throw$Result.Score[is.na(throw$Result.Score)]
<-mean(throw$Result.Score,na.rm = TRUE)
throw(is.na(throw$Result.Score))

```{r}
library(dplyr)
```


8. Form at least 5 questions by yourself to explore your data set and to get 
useful 
insights from that data. Apply appropriate data manipulation commands to achieve 
the 
same

a. Get the details of all the players from USA.

```{r}
filter(throw,Nationality=="USA")
```

b. Get the details of the players who have Rank 50

```{r}
filter(throw,Rank==50)
```

c. Find players who played in Salinas, CA (USA) and had a Result Score over 1280.

```{r}
filter(throw,Venue=="Salinas, CA (USA)",ResultScore>1280)
```

d. Get the DOB and Name of the competitor from EST

```{r}
throw%>%filter(Nationality=="EST")%>%select(Competitor, Nationality, Mark, DOB)
```

e. Find the maximum Mark scored in each nation.

```{r}
throw%>%group_by(Nationality) %>%
summarise(Max_Mark=max(Mark))%>%arrange(desc(Max_Mark))
```


* * *
## Conclusion:
Using the dataset of  Discus throw under Men, we learned about how to 
clean a dataset according to our needs.



