---
title: 'Lab 10: Healthcare Analytics'
author: "Loura Shiny M"
date: "11/11/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Healthcare Analytics

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to perform descriptive and predictive 
analytics from a suitable dataset chosen in the healthcare domain. 
The dataset can be either for regression or classification. 

## To study / implement:

From this Lab exercise we will be learning about Healthcare Anaylsis in R. 


## Dataset Description:

This dataset Life Expectancy (WHO) is a dataset the is collected from the
World Helath Organization (WHO) and United Nations (UN) websits. The dataset 
contains information, factors that influence the life expectancy of a person. 
Immunization related factors, Mortality factors, Economical factor, Social factors, 
like food habbits, etc was consider through the different columns such as 
bmi, GDP, population, etc.

https://www.kaggle.com/kumarajarshi/life-expectancy-who


## Objectives:

The objective her is to perform descriptive and predictive analytics for the 
chosen dataset - Life Expectancy (WHO). The descriptive analysis will give a 
summary of the dataset and answer the questions asked in that section
and using linear regression we will do predictive analysis and see if the 
Life Expectancy can be predicted from the given factors. 

------------------------------------------------------------------------


## Pre-Processing:

## Sequence of Steps:


### Importing required libraries 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(caret)
library(datarium)
```


### Importing Dataset

```{r}
Life_Expectancy = read.csv("Life Expectancy Data.csv")
```


### View structure of dataset 

```{r}
str(Life_Expectancy)
```


### Perform Data Clenaing


### Removing missing values 

```{r}
Life_Expectancy = na.omit(Life_Expectancy)
```


### Summary of the dataset 

```{r}
summary(Life_Expectancy)
```


## Descriptive Analysis:

##### Range of Adult Mortality 

```{r}
range(Life_Expectancy$Adult.Mortality)
```

The range of the Adult Mortality ranges from 1 to 723


##### Histogram of Income composition of resources

```{r}
hist(Life_Expectancy$Income.composition.of.resources,main="Income",xlab="Income Composition")
```


##### Show in a pie chart the composition of the status of the countries taken in the dataset

```{r}
y=table(Life_Expectancy$Status)
pie(y,col=c("blue","yellow"),main="Status")
```

The pie chart (pictorial representation) of the division of status of the countries 
taken into consideration for the dataset/


##### Boxplot of  Total expenditure

```{r}
boxplot(Life_Expectancy$Total.expenditure,mai=" Total.expenditure")
```

Box plots shows that there is a few outlier points detected.


##### Barplot to show the distribution of data collected in terms of years

```{r}
x=table(Life_Expectancy$Year)
barplot(x,main="Data Distribution - Year", xlab="Year", ylab="Frequency")
```


##### Plotting the Life Expectancy vs Schooling

```{r}
plot(Life_Expectancy$Life.expectancy,Life_Expectancy$Schooling,xlab="Life Expectancy",ylab="Schooling")
```


## Predictive Analysis:

### Model Building

Now to perform linear regression we need to remove the SPORT column. Thus, 

```{r}
Life_Expectancy_data=Life_Expectancy[,-c(1,3)]
str(Life_Expectancy_data)
```


##### Train the data considering TOTAL from the dataset

```{r}
set.seed(123)
train_samples <- Life_Expectancy_data$Life.expectancy   %>%
createDataPartition(p=0.8,list=FALSE)
```


##### Split the data into training and testing data

```{r}
train <- Life_Expectancy[train_samples,]
test <- Life_Expectancy[-train_samples,]
```


##### Build Linear model for the TOTAL and other columns in the datset. 
We need to see which attribute has best p-values


```{r}
Life_Expectancy_model<- lm(Life.expectancy ~.,data=Life_Expectancy_data)
summary(Life_Expectancy_model)
```


From summary we can see that Adjusted R-squared values is close to 1.
This is good as it indicates that a large propotion of the variability in the 
outcome has been explained by the regression model. Plus, 3 *'s 
(pr lesser than 0.05 ) for the columns Year, Adult.Mortality, infant.deaths, BMI,
under.five.deaths, HIV.AIDS, Income.composition.of.resources and Schooling 
show that we can use them to build linear model. 


Now, considering BMI and Life.expectancy for our model.

##### Outlier Detection

```{r}
boxplot(Life_Expectancy_data$BMI,ylab="BMI")
```

No, outlier is detected from boxplot.


###Plotting models

```{r}
hist(Life_Expectancy_model$residuals)
```


```{r}
qqnorm(Life_Expectancy_model$residuals,ylab = "Residuals")
qqline(Life_Expectancy_model$residuals)
```


```{r}
par(mar=c(2,2,2,2))
ggplot(Life_Expectancy_data,aes(x=BMI,y=Life.expectancy))+geom_point()+geom_smooth(method='lm',se=FALSE)
```


### Performance analysis:

##### Predicting the test

```{r}
lp<-Life_Expectancy_model%>%predict(test)
```


##### Finding RMSE

```{r}
RMSE <- RMSE(lp,test$Life.expectancy)
RMSE
```

The RMSE value is near to 0 which implese that a good outcome for the model.


##### Finding R2

```{r}
R2 <- R2(lp,test$Life.expectancy)
R2
```

The R2 value is great. This means the model output will be better.


##### Correlation Analysis

```{r}
cor(Life_Expectancy_data$BMI,Life_Expectancy_data$Life.expectancy)
```

Positive value indicates that the that there is positive correlation between 
BMI and Life.expectancy, i.e., they are directly  propositional to each other. 


##### Linear Model between TOTAL and STR

```{r}
model <- lm(Life.expectancy~BMI,data=Life_Expectancy_data)
model
```


```{r}
summary(model)
```


```{r}
sigma(model) 
```


```{r}
sigma(model)*100/mean(Life_Expectancy_data$Life.expectancy)  #percentage error
```

The percentage of error is 10%.


```{r}
confint(model)
```


##### Finding the confidence of BMI

```{r}
BMI=25
sdt <- data.frame(BMI)
conf<- predict(model,sdt,level = .95,interval = "confidence")
conf
```


##### Finding the prediction of STR

```{r}
lp <- predict(model,sdt,level = .95,interval = "prediction")
lp
```


##### Plotting model

```{r}
plot(model,1)
```


------------------------------------------------------------------------


## Results and Discussion: 

From descriptive and predictive analysis we got the desired results. 
In descriptive analysis, we can visualize the data from according to one's wish 
using different graphs. Using linear regression, we can predict the values of 
Life.expectancy from all the attributes as all of them had 3 *'s. 
And the example with BMI was shown.


* * *
## Conclusion:
Using the dataset Life Expectancy (WHO) we learnt about Descriptive and 
Predictive Analysis in R by performing them on the dataset.


## References:

https://pestleanalysis.com/descriptive-analysis/
https://machinelearningmastery.com/classification-versus-regression-in-machine-learning/
https://machinelearningmastery.com/machine-learning-in-r-step-by-step/
s
