---
title: 'Lab 8: Data Regression in R'
author: "Loura Shiny M"
date: "07/10/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Data Regression in R

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to learn about and execute the different 
problems under regression in R. 

## To study / implement:

From this Lab exercise we will be learning about Data Regression in R.


## Dataset Description:

The dataset weatherHistory is taken from Kaggle. It is a dataset meant to work on 
regression analysis. It includes a hourly/daily summary for Szeged, Hungary area, 
between 2006 and 2016. It contanins attributes such as time, summary, precipType,
temperature, humidity, windSpeed, pressure, etc. 

Dataset link: https://www.kaggle.com/budincsevity/szeged-weather


------------------------------------------------------------------------

## Questions

------------------------------------------------------------------------


Importing required libraries 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(caret)
library(datarium)
```


Importing Dataset

```{r}
weather = read.csv("weatherHistory.csv")
weather = na.omit(weather)
weather=head(weather,1000) #to get a preview of the dataset
weather=subset(weather, select = -c(Loud.Cover,Formatted.Date,Summary,Precip.Type,Daily.Summary) )
```

Structure of dataset

```{r}
str(weather)
```


Summary of the dataset weatherHistory

```{r}
summary(weather)
```


The task for this dataset is to predict the apparent temperature. Thus, first 
train the model according to the apparent temperature. 

```{r}
set.seed(123)
train_samples <- weather$Apparent.Temperature..C. %>%
  createDataPartition(p=0.8,list=FALSE)
```

```{r}
head(train_samples)
```

Now assign train and test datas

```{r}
train <- weather[train_samples,]
test <- weather[-train_samples,]
```


Now building linear model for apparent temperature and the other attributes

```{r}
wl <- lm(Apparent.Temperature..C.~.,data=weather)
summary(wl)
```

Inference: Residual median is equal to 1. As p-values of Humidity and Temperature
is less than 0.05, we can predict the values of apparent temperature from 
those values. Adjusted R- squared is 0.9889 which is equal to 1 tells that the 
model can work as a regression model.

Predict the test

```{r}
pred <- wl %>%
  predict(test)
```


Finding RMSE

```{r}
r<- RMSE(pred,test$Apparent.Temperature..C.)
r
```


Finding R2

```{r}
r2 <- R2(pred,test$Apparent.Temperature..C.)
r2
```

Histogram of the model 

```{r}
hist(wl$residuals)
```

Qqnorm

```{r}
qqnorm(wl$residuals,ylab="Residuals")
qqline(wl$residuals)
```

```{r}
par(mar=c(2,2,2,2))
ggplot(weather,aes(x=Humidity,y=Apparent.Temperature..C.))+geom_point()+geom_smooth(method='lm',se=FALSE)
```


```{r}
cor(weather$Humidity,weather$Apparent.Temperature..C.)
```
Since we get a negative value we can say that Apparent temperature and humidity 
are inversely propositional to each other. 


* * *
## Conclusion:
Using the dataset weatherHistory we learnt about Regression in R.