---
title: 'Lab 6: Missing Data in R'
author: "Loura Shiny M"
date: "23/09/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Missing Data in R

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to learn and execute about the different 
concepts under missing data. We'll see how one can view the amount of NA data
in today's Lab

## To study / implement:

From this Lab exercise we will be learning about Missing Data in R.


## Dataset Description:

The dataset used is mtcars (Motor Trend Car Road Tests). It is available in .
The data was extracted from the 1974 Motor Trend US magazine, and comprises 
fuel consumption and 10 aspects of automobile design and performance for
32 automobiles (1973–74 models). It consists of many attributes like mpg, cyl,
hp, drat, wt, etc


------------------------------------------------------------------------

## Questions


Loading in the dataset:

```{r}
data <- mtcars
data
```

Summary of data
```{r}
summary(data)
```


Install required package

```{r}
#install.packages("missForest")
```

```{r}
#install.packages("mice")
```

```{r}
#install.packages("VIM")
```

Load library

```{r}
library(missForest)
```

```{r}
library(mice) 
```

```{r}
library(VIM)
```


```{r}
set.seed(10)  #random number generated 
```

------------------------------------------------------------------------

## With NA probability 10% :

```{r}
datan <- prodNA(data, noNA = 0.1)
datan
```

```{r}
md.pattern(datan)
#shows the pattern of occurences of NA values
```

```{r}
datamiss = aggr(datan, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, 
labels=names(datan), cex.axis=.7, gap=3, ylab=c("Proportion","Pattern"))
```

```{r}
mice_imputes = mice(datan)
```

```{r}
mice_imputes$method
```


```{r}
Imputed_data=complete(mice_imputes)
datan
Imputed_data
```


------------------------------------------------------------------------

## With NA probability 20% :

```{r}
datan <- prodNA(data, noNA = 0.2)
datan
```

```{r}
md.pattern(datan)
#shows the pattern of occurences of NA values
```

```{r}
datamiss = aggr(datan, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, 
labels=names(datan), cex.axis=.7, gap=3, ylab=c("Proportion","Pattern"))
```

```{r}
mice_imputes = mice(datan)
```

```{r}
mice_imputes$method
```


```{r}
Imputed_data=complete(mice_imputes)
datan
Imputed_data
```


------------------------------------------------------------------------

## With NA probability 5% :

```{r}
datan <- prodNA(data, noNA = 0.05)
datan
```


```{r}
md.pattern(datan)
#shows the pattern of occurences of NA values
```

```{r}
datamiss = aggr(datan, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, 
labels=names(datan), cex.axis=.7, gap=3, ylab=c("Proportion","Pattern"))
```

```{r}
mice_imputes = mice(datan)
```

```{r}
mice_imputes$method
```


```{r}
Imputed_data=complete(mice_imputes)
datan
Imputed_data
```

* * *
## Conclusion:
Using the dataset mtcars we learnt about Missing Data in R and how we can view 
patterns of missing data. 

