---
title: "Lab 2: Lists and Dataframes"
author: "Loura Shiny M"
date: "15/08/2021"
output:
  pdf_document: default
  html_document: default
---

------------------------------------------------------------------------

##### Name   : Loura Shiny M

##### Reg.No : 19BCE1524

##### Title  : Lists and Dataframes 

------------------------------------------------------------------------

## Aim

The aim of this lab experiment is to create an employee list and dataframe and execute the instructions given in each question.

## To study / implement:

From this Lab exercise we will be learning about lists and dataframes.

## Dataset Description

This is a simple employee dataset. It will contain details of employee such as their ID, name, designation, company name, date of birth, contact number, mail ID, etc.

------------------------------------------------------------------------

## Questions

1.Create a list for an employee with the following fields:EmpId, Name, Designation, CompanyName

```{r}
employee<-list(EmpId=c("B99100"), Name=c("Amy Santiago"), Designation=c("Manager"), CompanyName=c("Brookyln Nine Nine"));
employee
```

2. Add the fields viz., DateOfBirth, ContactNumber and MailID to the 
above list.

```{r}
employee<-append(employee,list(DOB=as.Date("1985-05-05"),ContactNumber=c("9876543210"),MailID=c('amysantiago@gmail.com')))
employee
```

3. Print the total number of fields and the field names.

```{r}
str(employee)
```
```{r}
length(employee)
```
```{r}
names(employee)
```

4. Create a data frame with the above fieldnames (5 different variables for EmpId, Name, Designation, Company, DOB) for 5 employees.

```{r}
EmpId=c("B991001","B991002","B991003","B991004","B991005")
Name=c("Amy Santiago","Jake Peralta","Terry Jeffords","Rosa Diaz","Charles Boyle")
Designation=c("Manager","Engineer","Regional Head","Engineer","Technician")
Company=c("Brooklyn Nine Nine","Brooklyn Nine Nine","Brooklyn Nine Nine","Brooklyn Nine Nine","Brooklyn Nine Nine")
d1="1985-05-05"
d1<-as.Date(d1)
d2="1985-10-15"
d2<-as.Date(d2)
d3="1975-09-23"
d3<-as.Date(d3)
d4="1990-01-12"
d4<-as.Date(d4)
d5="1983-08-09"
d5<-as.Date(d5)
DOB=c(d1,d2,d3,d4,d5)
employeedf=data.frame(EmpId, Name, Designation, Company, DOB)
employeedf
```

5. Add the ContactNumber and MailID to this already created data frame.

```{r}
employeedf$ContactNumber<-c("9876543210","1234567890","3786297501","2491823018","91827362071")
employeedf
```
```{r}
employeedf$MailID<-c("amysantiago@gmail.com","jakeperalta@gmail.com","teryyjefords@gmail.com","rosadiaz@gmail.com","charlesboyle@gmail.com")
employeedf
```

6. Add the details of employee 3 without contact number and employee4
without mentioning mailID.
```{r}
employeedf[3,6]<-NA
employeedf[4,7]<-NA
employeedf
```

7. Display the details of the employees after removing the missing values.

```{r}
employeedf1=employeedf[complete.cases(employeedf), ]
employeedf1
```

8. Display the details of employees who celebrate birthday in the month of August.

```{r}
subset(employeedf,months(as.Date(employeedf$DOB))=="August")
```

9. Display the details of employees whose designation is “Manager”.
```{r}
employeedf[employeedf$Designation=="Manager", ] 
```

10.Calculate the age of all the employees (as on August 12, 2021) and include it as an additional column in the same data frame.

```{r}
library(eeptools)
```

```{r}
Age=c(floor(age_calc(employeedf$DOB,units="years"))) 
employeedf=cbind(employeedf,Age) 
employeedf
```

